\( \tool \): JavaScript IR-based Semantics Extraction Toolchain

### 2020-03-28 ase20/abstract.tex
[ASE] Revised abstract / intro

\begin{abstract}
JavaScript was initially designed for client-side programming in web browsers,
but its engine is now embedded in various kinds of host software.  Despite the
popularity, since the JavaScript semantics is complex especially due to its
dynamic nature, understanding and reasoning about JavaScript programs are
challenging tasks.  Thus, researchers have proposed several attempts to define
the formal semantics of JavaScript based on ECMAScript, the official JavaScript
specification.  However, the existing approaches are manual, labor-intensive,
and error-prone and all formal semantics target former versions of ECMAScript
5.1 (ES5.1).  It is critical in understanding modern JavaScript language
features introduced after ECMAScript 6 (ES6). Moreover, ECMAScript has been
annually released since 2015, which made already five updates after ES5.1.

To alleviate the problem, we propose \( \tool \), a JavaScript IR-based
Semantics Extraction Toolchain. It is the first tool that \textit{automatically
synthesizes} parsers and AST-IR translators directly from a given ECMAScript.
For syntax, we develop a parser generation technique with \textit{lookahead
parsing} for \( \bnfes \), a variant of the extended BNF used in ECMAScript.
For semantics, \( \tool \) synthesizes AST-IR translators using \textit{forward
compatible} rule-based compilation. Each \textit{compile rule} describes how to
convert each step of abstract algorithms written in structured natural language
into \( \ires \), an Intermediate Representation that we designed for
ECMAScript. For the five most recent ECMAScript versions, \( \tool \) succeeds
to synthesize parsers for all versions, and to compile \inred{XX.XX}\% of
algorithm steps in average. We manually complete the core part of formal
semantics of the next version of ECMAScript (ES11) based on the extracted
semantics and it passed all \inred{XX,XXX} applicable tests. Based on the formal
semantics, we found \inred{X} specification errors in ES11 and all of them were
confirmed by TC 39.  Furthermore, we also successfully applied \( \tool \) to
\inred{X} feature proposals in Stage 3.
\end{abstract}

>>>>

\begin{abstract}
JavaScript was initially designed for client-side programming in web browsers,
but its engine is now embedded in various kinds of host software.  Despite the
popularity, since the JavaScript semantics is complex especially due to its
dynamic nature, understanding and reasoning about JavaScript programs are
challenging tasks.  Thus, researchers have proposed several attempts to define
the formal semantics of JavaScript based on ECMAScript, the official JavaScript
specification.  However, the existing approaches are manual, labor-intensive,
and error-prone and all formal semantics targeted ECMAScript 5.1 (ES5.1, 2011)
or more former versions.  It is critical in understanding modern JavaScript
language features introduced by ECMAScript 6 (ES6, 2015).  Moreover, ECMAScript
has been annually released since 2015, which made already five updates after
ES5.1.

To alleviate the problem, we propose \( \tool \), a JavaScript IR-based
Semantics Extraction Toolchain.  It is the first tool that \textit{automatically
synthesizes} parsers and AST-IR translators directly from a given ECMAScript.
For syntax, we develop a parser generation technique with \textit{lookahead
parsing} for \( \bnfes \), a variant of the extended BNF used in ECMAScript.
For semantics, \( \tool \) synthesizes AST-IR translators using \textit{forward
compatible} rule-based compilation.  Each \textit{compile rule} describes how to
convert each step of abstract algorithms written in structured natural language
into \( \ires \), an Intermediate Representation that we designed for
ECMAScript.  For the five most recent ECMAScript versions, \( \tool \) succeeds
to synthesize parsers for all versions, and to compile \inred{XX.XX}\% of
algorithm steps in average.  We manually complete the core part of formal
semantics of the draft of the next ECMAScript (ES11, 2020) based on the
extracted semantics and it passed all \inred{XX,XXX} applicable tests.  Based on
the formal semantics, we found \inred{X} specification errors in ES11 and all of
them were confirmed by TC 39.  Furthermore, we also successfully applied \(
\tool \) to \inred{X} feature proposals in Stage 3 to check the forward
compatibility of \( \tool \).
\end{abstract}

### 2020-03-28 ase20/intro.tex
[ASE] Revised abstract / intro

\section{Introduction}
JavaScript is one of the most widely used programming languages not only for client-side
but also for server-side programming~\cite{nodejs, meanjs}
and even for small embedded systems~\cite{espruino, tessel2}.
It is the top-ranked language used in active GitHub
repositories\footnote{https://githut.info/}, and \#7 in the TIOBE
Programming Community index\footnote{https://www.tiobe.com/tiobe-index/}.
According to W3Techs\footnote{https://w3techs.com/technologies/details/cp-javascript/all/all},
95.2\% of websites use JavaScript as their client-side programming language.

Despite its popularity, JavaScript developers often suffer from its intricate semantics,
which may cause unexpected behavior.  For example, calling the following JavaScript
function seems to always return \( \code{false} \):
\begin{lstlisting}[style=myJSstyle]
    function f(x) { return x == !x; }
\end{lstlisting}
Unfortunately, it returns \( \code{true} \) when its argument is an empty array
\( \code{[]} \).  To correctly understand and reason about such complex
behavior, the formal semantics of JavaScript is necessary.

\begin{figure}
  \centering
  \includegraphics[width=0.48\textwidth]{img/existing.png}
\vspace*{-2em}
  \caption{Existing approaches: Manual extraction of IR-based semantics for ECMAScript}
  \label{fig:existing}
\vspace*{-1em}
\end{figure}

Researchers have defined various JavaScript formal
semantics~\cite{aplas08,lambdajs,kjs,javert} and developed several static
analyzers on top of them~\cite{jsai,tajs,wala,safe}.  As illustrated
in Figure~\ref{fig:existing}, they defined their own Intermediate
Representations (IRs) to represent JavaScript program semantics.  They
built parsers to construct Abstract Syntax Trees (ASTs) from
programs and AST-IR translators to convert ASTs
to their own IRs.  This approach helps researchers focus on IRs
without worrying about diverse and enormous features of JavaScript in
developing new techniques for program analysis and formal verification.
We dub such a methodology that manipulates language semantics by
translating given programs to IRs that are suitable for analysis and
verification \textit{IR-based semantics extraction}.

However, existing approaches to JavaScript IR-based semantics
extraction have difficulties in manipulating recent JavaScript
programs, because they all extract the semantics \textit{manually}.
The JavaScript syntax and semantics are described in the ECMAScript
specification, written in English in a clearly structured and
organized manner.  Unfortunately, manually implementing parsers and
translators for JavaScript from the specification is tedious,
labor-intensive, and error-prone especially because the language is
huge.  The most recent specification, ECMAScript 2019, has 798 pages.

Moreover, manual extraction does not scale for frequent updates of the
specification.  Most existing formal semantics and analyzers of
JavaScript support only ECMAScript 5.1~\cite{ecma5} released in June 2011.
Focusing on ECMAScript 5.1 was reasonable until 2014, because the
specification was rarely updated.  However, in late 2014, The Ecma
Technical Committee 39 (TC39)~\cite{tc39} decided to release the
specification annually.  Currently, no formal semantics
nor analyzers for JavaScript correctly deal with recent ECMAScript
features such as lexical binding via \( \code{let} \), the spread
\( \code{...} \) operator, classes, the \( \code{for-of} \) operator,
the \( \code{async} \) functions, and generators.

In this paper, we propose JavaScript IR-based Semantics Extraction
Toolchain (\( \tool \)), which \textit{automatically} extracts JavaScript
syntax and semantics from the ECMAScript specification.  It
automatically generates a JavaScript parser from the syntax defined in
the ECMAScript specification.  It also automatically extracts the
formal semantics of JavaScript with given \textit{compile rules}.
Each compile rule describes how abstract algorithms in the
specification written in a natural language are converted into
functions of our intermediate representation \( \ires \).  In
addition, to save the labor of defining such compile rules, we also
present a \textsf{Fault Localizer}, which suggests new
compile rules based on fault localization of not yet convertible
algorithms.  It helps us deal with changed or newly
introduced algorithms defined in new versions of the ECMAScript
specification as well.

The main contribution of this paper is \( \tool \):
\begin{itemize}[leftmargin=0.5cm]
\item We successfully applied \( \tool \) to ECMAScript 2020, the next
ECMAScript version scheduled to be released in June 2020.  The
automatically extracted semantics passed the official conformance test
suite, Test262~\cite{test262}; out of 19,282 tests for
ECMAScript 2020 it passed 19,220 tests.
\item \( \tool \) is also applicable to new language features proposed
for a future release to aid language development in a modular way.
Because ECMAScript is an open standard language, various proposals are
available with their own specification changes and tests.  As a case
study, we applied \( \tool \) to such a proposal, during which the tool
found some errors in the proposal.  The reported errors were all confirmed
and the tool passed all the tests.
\item Moreover, since the extraction mechanism of \( \tool \) is fully
automatic, we can use it to find possible specification errors.  We
indeed found 3 specification errors in ECMAScript 2020.  They were all
confirmed by TC39 and will be fixed in the next release.
\end{itemize}

>>>>

\section{Introduction}
JavaScript is one of the most widely used programming languages not only for
client-side but also for server-side programming~\cite{nodejs, meanjs} and even
for small embedded systems~\cite{espruino, tessel2}.  It is the top-ranked
language used in active GitHub repositories\footnote{https://githut.info/}, and
\#7 in the TIOBE Programming Community
index\footnote{https://www.tiobe.com/tiobe-index/}.  According to
W3Techs\footnote{https://w3techs.com/technologies/details/cp-javascript/all/all},
95.0\% of websites use JavaScript as their client-side programming language.

Despite its popularity, JavaScript developers often suffer from its intricate
semantics, which may cause unexpected behaviors.  For example, calling the
following JavaScript function seems to always return \( \code{false} \):
\begin{lstlisting}[style=myJSstyle]
function f(x) { return x == !x; }
\end{lstlisting}
Unfortunately, it returns \( \code{true} \) when its argument
is an empty array \( \code{[]} \).  To correctly understand and reason about
such a complex behavior, the formal semantics of JavaScript is necessary.

\begin{figure}
  \centering
  \includegraphics[width=0.48\textwidth]{img/existing.pdf}
\vspace*{-2em}
  \caption{Existing approaches: Manually built parsers and AST-IR
  translators for JavaScript IR-based semantics}
  \label{fig:existing}
\vspace*{-1em}
\end{figure}

Researchers have defined various JavaScript formal
semantics~\cite{aplas08,lambdajs,kjs,javert} suitable for static
analysis~\cite{jsai,tajs,wala,safe} and formal verification~\cite{javert} by
referring to ECMAScript.  ECMAScript is the official specification that
describes JavaScript syntax using a variant of the extended BNF (EBNF) notation,
and its semantics using abstract algorithms written in English in a clear and
structured manner.  \textit{IR-based semantics extraction} is a typical way to
define the formal semantics by building a compiler from programs to Intermediate
Representations (IRs) to indirectly represent the semantics of given programs.
As illustrated in Figure~\ref{fig:existing}, the compiler consists of a parser
that constructs Abstract Syntax Trees (ASTs) of given JavaScript programs, and
an AST-IR translator that converts ASTs to their own IRs.  This approach helps
researchers focus on IRs without worrying about diverse and enormous features of
JavaScript in developing new techniques for static analysis and formal
verification.

However, to the best of our knowledge, all existing approaches to JavaScript
IR-based semantics extraction \textit{manually} build parsers and translators.
Although it was reasonable approach until ECMAScript 5.1 (ES5.1,
2011)~\cite{es5}, the manual approach is too tedious, labor-intensive, and
error-prone to deal with the large size of modern JavaScript since ECMAScript 6
(ES6, 2015)~\cite{es6}. ES6 introduced lots of significantly new features such
as lexical binding via \( \code{let} \), the spread \( \code{...} \) operator,
classes, the \( \code{for-of} \) operator, the \( \code{async} \) functions, and
generators.  KJS~\cite{kjs} is one of formal semantics of ECMAScript 5.1 (ES5.1,
2011) and it is defined on top of \( \kframework \), which is a framework for
defining language semantics.  According to an author of KJS, it took
\textit{four months} to implement AST-IR translator for 1,370 steps out of 2,932
steps in 368 abstract algorithms. However, the most recent version of ECMAScript
(ES10, 2019)~\cite{es10} has 1,731 abstract algorithms consisting of 10,068
steps. Thus, it is not scalable to build AST-IR translator for modern JavaScript
and no formal semantics exists for ES6 to ES10.

Moreover, JavaScript syntax and semantics are annually updated.  Until
ECMAScript 5.1, JavaScript was a stable language because the specification was
rarely updated.  However, The Ecma Technical Committee 39 (TC39)~\cite{tc39}
decided to release the specification annually in late 2014.  After this official
announcement, several syntax features and roughly 1,000 to 3,000 steps of
abstract algorithms are modified or newly added in the specification every year.
To deal with these frequent and massive updates of ECMAScript, researchers
should manually update parsers and AST-IR translators by putting a lot of
effort.

To alleviate this problem, we propose a technique to \textit{automatically
synthesize} parsers and AST-IR translators directly from ECMAScript with
\textit{forward compatibility}. There are several technical challenges in
synthesizing parsers and translators. For syntax, ECMAScript utilize their own
special variant of the EBNF with parametric non-terminals, conditional
alternatives, and various special terminal symbols.  Thus, no existing parser
generation technique is directly applicable for this variant.  Moreover,
JavaScript provides automatic semicolon insertion in its parsing algorithm with
several complex rules not in a lexer. For semantics, abstract algorithms in
ECMAScript are written in English.  Besides, general and forward compatible
representation of abstract algorithms is necessary for future version of
ECMAScript.

Our contribution is \( \tool \), a JavaScript IR-based Semantics Extraction
Toolchain:
\begin{itemize}[leftmargin=0.5cm]
  \item \( \tool \) is the \textit{first} tool that can automatically synthesize
    parsers and AST-IR translators for modern JavaScript directly from
    ECMAScript.  For syntax, we formalize the variant of EBNF as \( \bnfes \)
    for the first time, and propose a parser generation technique with
    \textit{lookahead parsing} for \( \bnfes \) with automatic semicolon
    insertion. For Semantics, synthesis of AST-IR translators are semi-automatic
    assisted by \textit{compile rules} because of the difficulty in dealing with
    the semantics written in a natural language.  Compile rules describe how
    each step of abstract algorithms is converted into our intermediate
    representation \( \ires \) designed for ECMAScript. We evaluated \( \tool \)
    for the five most recent ECMAScript versions including the draft of the next
    version (ES6 to ES11).  We succeed to generate parsers for all versions, and
    to automatically cover \inred{XX.X}\% of steps in abstract algorithms in
    average.
  \item Based on \( \tool \), we define the first IR-based formal semantics of
    modern JavaScript.  The semantics targets the core part of the draft of next
    version of ECMAScript (ES11), and our tool successfully synthesizes its
    parser and AST-IR translator for \inred{X,XXX} out of \inred{X,XXX}
    algorithm steps.  We manually complete the missing parts of the AST-IR
    translator (\inred{XXX} steps) in \inred{XX} days.  To check the correctness
    of the formal semantics, we evaluated Test262~\cite{test262}, the official
    conformance test suite, and the semantics passed all \inred{XX,XXX}
    applicable tests.  Moreover, we found \inred{X} specification errors in ES11
    based on the formal semantics.  They were all confirmed by TC39 and will be
    fixed in the next release.
  \item \( \tool \) is also forward compatible with new language features
    proposed for future ECMAScript specifications.  TC39 manages proposals of
    new language features in an open source
    project\footnote{https://github.com/tc39/proposals}.  There are five stages
    of proposals (from Stage 0 to Stage 4).  All proposals in Stage 4 (Finished)
    are applied in the draft of the next ECMAScript, Stage 3 (Candidate) is the
    last stage before Stage 3.  For evaluation of forward compatibility of \(
    \tool \), we applied our tool to all \inred{XX} Stage 3 proposals. It
    succeeded to synthesize parsers for all proposals, and to convert \inred{XX}
    out of \inred{XX} algorithm steps for all proposals. After completing all
    missing parts, each semantics passed all \inred{XX} tests.
\end{itemize}
